---
title: "CBC-01 Seropositive proportions"
author: "Mun-Gwan"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_document:
    theme: readable
    toc: true
    toc_float: true
    toc_depth: 3
    fig_caption: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../results/sero_positive_by_sd") })
---

<style>
p.caption {
  font-size: 1em;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	dev = c('png', 'pdf'),
	fig.path = '../results/sero_positive_by_sd/figures/',  # save figure in files
	results = 'asis'
)

library(tidyverse)
library(Useful2me)  # remotes::install_github("Rundmus/Useful2me-R_package", upgrade= "never")
library(knitr)
library(kableExtra)     # kable_styling
library(UpSetR)
library(table1)

# library(janitor)     # tabyl, fisher.test

rm(list = ls())


###  File names

source("options_master_trans_serology.R")


fn <- list(
	i = list(                               #  input
		b1 = data_file_path_batch1,
		b2 = data_file_path_batch2,
		binder = '../data/assay_information/bead_id_information/CBA5_information_js_matching.csv',
		umap = list(
			set1 = '../results/dim_reduction_plots/batch1/luminex/2020-06-22_dim_red_dbs_covid19_manuscript_batch1_cba_norm_v1_detectionAb_IgG_IgM_seropositivity.txt',
			set2 = '../results/dim_reduction_plots/batch2/luminex/2020-06-22_dim_red_dbs_covid19_manuscript_batch2_cba_norm_v1_detectionAb_IgG_IgM_seropositivity.txt'
		)
	)
)

#  Check all exists
stopifnot(all(file.exists(c('.', unlist(fn$i)))))


###  Load data

load(fn$i$b1)
bBs <- list(
	set1 = list(
		sinfo = sinfo,
		mfi = mfi,
		binder = binder
	)
)

tmp <- load(fn$i$b2)
bBs$set2 <- list(
	sinfo = sinfo,
	mfi = mfi,
	binder = binder
)

rm(list = tmp)


cbas <- read_csv(
	fn$i$binder,
	col_types = cols_only(
		# CBA_nr = col_character(),
		# Binder = col_character(),
		# Comment = col_character(),
		Bead_ID = col_character(),
		# Coupling_chemistry = col_character(),
		Name = col_character(),
		# Provider = col_character(),
		# `Edsberg Analysis` = col_character(),
		`In batch 1` = col_character(),
		`Batch 2 Analysis` = col_character(),
		Match_Batch_1 = col_character()
	)
)
stopifnot(anyDuplicated(cbas$Name) == 0)


##  CONSTANTS

SERONEG_PROP = 0.8    # assumed sero-negative proportion 
PREVALENCE = 0.1      # assumed the prevalence of Covid-19
```



----------------------------

###  Seropositive


```{r prop,results='markup'}

##  Only the binders in common

bBs$set1$binder <- cbas %>% 
	separate_rows(Match_Batch_1, sep = ", ") %>% 
	distinct(Match_Batch_1, .keep_all = T) %>% 
	drop_na(Match_Batch_1) %>% 
	dplyr::select(Name = Match_Batch_1, `Batch 2 Analysis`) %>% 
	left_join(bBs$set1$binder, ., by= "Name") 

common_in_set1 <- bBs$set1$binder %>% 
	filter(`Batch 2 Analysis` == "Include") %>% 
	pull(id)

bBs$set1$binder <- bBs$set1$binder %>% 
	filter(id %in% common_in_set1)

bBs$set1$mfi <- bBs$set1$mfi[, colnames(bBs$set1$mfi) %in% common_in_set1]
stopifnot(identical(colnames(bBs$set1$mfi), bBs$set1$binder$id))


bBs$set2$binder <- bBs$set2$binder %>% 
	left_join(cbas %>% dplyr::select(Name, `In batch 1`, `Batch 2 Analysis`), by= "Name")

# bBs$set2$mfi <- bBs$set2$mfi[, bBs$set2$binder$`In batch 1` == "Yes"] 
# bBs$set2$binder <- bBs$set2$binder %>% 
# 	filter(`In batch 1` == "Yes")
stopifnot(identical(colnames(bBs$set2$mfi), bBs$set2$binder$id))


##  Population, Positive and Negative only

bBs <- lapply(
	bBs,
	function(ea) {
		ea$sinfo <- ea$sinfo %>% 
			mutate(
				grp = case_when(
					grepl("CBC\\+", subj.id) ~ "Population",
					grepl("Positive", subj.id) ~ "Positive",
					grepl("Negative", subj.id) ~ "Negative",
					TRUE ~ NA_character_
				)
			)
		ea$mfi <- ea$mfi[!is.na(ea$sinfo$grp), ]
		ea$sinfo <- ea$sinfo %>% 
			drop_na(grp)
		return(ea)
	}
)

stopifnot(
	identical(rownames(bBs$set1$mfi), bBs$set1$sinfo$id),
	identical(rownames(bBs$set2$mfi), bBs$set2$sinfo$id)
)

##  Sero-positive or not (T or F) matrix per set and cutoff
posneg <- lapply(c("set1", "set2") %>% `names<-`(., .), function(iset) {
	# iset = "set1"
	
	#  No duplicated CBC+01 samples
	stopifnot(
		anyDuplicated(
			bBs[[iset]]$sinfo$subj.id[grepl("CBC\\+", bBs[[iset]]$sinfo$subj.id)]
		) == 0
	)
	
	subj.id <- bBs[[iset]]$sinfo$subj.id
	id <- bBs[[iset]]$sinfo$id
	grp <- bBs[[iset]]$sinfo$grp

	#  All population samples
	mfi_pop <- bBs[[iset]]$mfi[grp == "Population", ]
	
	cat("\n> ", iset, "\n")
	for(igrp in unique(grp)) {
		cat("\n* Dimension of", igrp, "samples\n")
		print(dim(bBs[[iset]]$mfi[grp == igrp, ]))
		
		if(igrp %in% c("Positive", "Negative")) {
			#  Known positive/negative control samples
			bBs[[iset]]$sinfo %>% 
				filter(grp == igrp) %>% 
				pull(id) %>% 
				sub("\\*[[:digit:]]*", "", .) %>% 
				table() %>% 
				print()
		}
	}

	#  Finding cutoff ---------------------------------------------------------
	
	mfi_pop <- mfi_pop %>%
		as_tibble()
	
	##  Based on peak value
	den <- mfi_pop %>%
		lapply(density)
	
	#  peak
	m <- sapply(den, function(x) x$x[which.max(x$y)])
	
	#  std.dev. of (SERONEG_PROP) around the peak
	s <- sapply(seq(m), function(ii) {
		df <- abs(mfi_pop[[ii]] - m[ii])
		qtl <- quantile(df, probs = c(SERONEG_PROP))
		sd(mfi_pop[[ii]][df <= qtl])
	})

	
	lapply(c(3, 6) %>% `names<-`(., .), function(isd) {            # isd <- 3
		#  Cutoff = peak + 'isd' * std.dev. of (SERONEG_PROP) around the peak
		cutoff <- m + isd * s
		
		#  Positve = TRUE, Negative = FALSE
		truefalse <- function(coff, mat) {
			rep(coff, each= nrow(mat)) %>%
				matrix(., nrow= nrow(mat)) %>%
				{mat > .} %>%
				as_tibble()
		}
		
		truefalse(cutoff, bBs[[iset]]$mfi) %>% 
			mutate(grp = grp, sample_id = id, subj.id = subj.id) %>% 
			dplyr::select(sample_id, subj.id, grp, everything())
	})
})	
```


###  Tables

*  95% confidence interval (CI) was computed with the assumption of random sample from Stockholm population and normal approximation. 
*  Postive predictive value (PPV) was computed assuming the prevalance of the disease was 10%.
*  During computation of Sensitivity, Specificity and PPV, repeated measured control samples were not taken into consideration. 


```{r table}
for(iset in names(posneg)) {
	
	for(isd in names(posneg[[iset]])) {
		out <- posneg[[iset]][[isd]] %>% 
			
			pivot_longer(-c(grp, sample_id, subj.id), names_to = "id") %>%
			group_by(grp, id) %>%
			summarise(
				Sero_pos = sum(value),        # sero-positive
				Non_reac = sum(!value)        # non-reactive (= sero-negative)
			) %>%
			mutate(
				#  Add binder info
				i_match = match(id, bBs[[iset]]$binder$id),
				Binder = bBs[[iset]]$binder$Binder[i_match],
				Name = bBs[[iset]]$binder$Name[i_match],
				det_ab = bBs[[iset]]$binder$detect_Ab[i_match],
				#  Positive proportion
				pos_prop = Sero_pos / (Sero_pos + Non_reac)
			) %>%
			dplyr::select(-i_match) %>% 
			arrange(det_ab) %>% 
			ungroup()

		#   Sens, Spec and PPV computation
		out1 <- out %>% 
			dplyr::select(-Sero_pos, -Non_reac) %>% 
			pivot_wider(names_from = grp, values_from = pos_prop) %>% 
			dplyr::select(
				id, 
				"Pos. prop." = Population, 
				Sensitivity = Positive, 
				Specificity = Negative
			) %>% 
			mutate(
				Specificity = 1 - Specificity,
				PPV = (Sensitivity * PREVALENCE) / 
					(Sensitivity * PREVALENCE + (1 - Specificity) * (1 - PREVALENCE))
			)
		
		out1 <- out1 %>% 
			mutate(
				#  P +- dev
				dev = sqrt(`Pos. prop.` * (1 - `Pos. prop.`) / sum(posneg[[iset]][[isd]]$grp == "Population")) * -qnorm(0.025),
				#  lower bound of CI
				lower = `Pos. prop.` - dev,
				lower = if_else(lower < 0, 0, lower),
				#  upper bound of CI
				upper = `Pos. prop.` + dev
			) %>% 
			#  Change to "0.00%" character
			mutate_at(
				c("Pos. prop.", "lower", "upper"),
				function(x) {
					if_else(
						is.na(x),
						"",
						(x * 100) %>% 
							round(., 2) %>% 
							format(width = 5) %>% 
							paste0(., "%")
					)
				}
			) %>% 
			mutate(
				`95% CI` = paste0(lower, "-", upper)
			) %>% 	
			#  Change to "00%" character
			mutate_if(
				is.numeric,
				function(x) {
					if_else(
						is.na(x),
						"",
						(x * 100) %>% 
							round() %>% 
							format(width = 5) %>% 
							paste0(., "%")
					)
				}
			)
			
		out %>% 
			filter(grp == "Population") %>% 
			left_join(out1, by= "id") %>% 
			dplyr::select(
				"Detect Ab" = det_ab, 
				Binder, 
				Name,
				`Pos. prop.`, 
				`95% CI`,
				'Pos.' = Sero_pos, 
				'Neg.' = Non_reac,
				'Sens' = Sensitivity,
				'Spec' = Specificity,
				PPV
			) %>% 
			kable(caption = paste0("Seropositves in ", sub("set", "Set ", iset), 
														 " (Cutoff ", isd, " S.D.)")) %>% 
			kable_styling(full_width = F) %>% 
			print()
	}
}
```

-------------------------------------

###  UpSet plots - All SPK,RBD,NCP binders


```{r upsetplot, fig.align='center'}


for(isd in names(posneg[[1]])) {
	for(iset in names(posneg)) {
		
		for(i_det_ab in c("IgG", "IgM")) {
			
			cat("\n#### ", sub("set", "Study set ", iset), "-", i_det_ab, " (Cutoff ", isd, " S.D.)\n")

			sel_abs <- (bBs[[iset]]$binder$detect_Ab == i_det_ab) & 
				(grepl("SPK|NCP|RBD", bBs[[iset]]$binder$Name))
			
			sel_abs_names <- bBs[[iset]]$binder %>% 
				dplyr::select(id, Name, detect_Ab) %>% 
				filter(sel_abs) %>% 
				unite(Name, Name, detect_Ab, sep = " - ") 
			sorted_sel_abs_names <- sel_abs_names %>% 
				separate(Name, c("Name1", "Name2"), sep= "_") %>% 
				arrange(Name1, desc(Name2)) %>% 
				unite(Name, Name1, Name2) %>% 
				pull(Name)

			# rtio <- (0.4 - 0.25) / (10 - 4) * (length(sorted_sel_abs_names) - 4) + 0.25
			
			posneg[[iset]][[isd]] %>% 
				filter(grp == "Population") %>% 
				select_at(sel_abs_names$id) %>% 
				`colnames<-`(sel_abs_names$Name) %>% 
				mutate_all(function(x) if_else(x, 1L, 0L)) %>% 
				as.data.frame() %>% 
				# as.list() %>% 
				# lapply(which) %>% 
				# 
				upset(nsets = 10, 
							sets = sorted_sel_abs_names,
							nintersects = 40,
							order.by = "degree",
							keep.order = T,
							text.scale = c(1, 1.5, 1, 1, 1.2, 1),
							# mb.ratio = c(1 - rtio, rtio),
							mainbar.y.label = "Number of seropositive samples [N]") %>% 
				print()
			cat("\n")
		}
	}
}


```

-------------------------------------

###  UpSet plots - SPK_02, RBD_01, NCP_01


```{r upset_sets, fig.align='center'}


for(isd in names(posneg[[1]])) {
	for(iset in names(posneg)) {
		
		cat("\n#### ", sub("set", "Study set ", iset), " (Cutoff ", isd, " S.D.)\n")
		
		sel_abs <- bBs[[iset]]$binder$Name %in% c("SPK_02", "NCP_01", "RBD_01")
		
		sel_abs_names <- bBs[[iset]]$binder %>% 
			dplyr::select(id, Name, detect_Ab) %>% 
			filter(sel_abs) %>% 
			unite(Name, Name, detect_Ab, sep = " - ") 
		
		# rtio <- (0.4 - 0.25) / (10 - 4) * (length(sorted_sel_abs_names) - 4) + 0.25
		
		posneg[[iset]][[isd]] %>% 
			filter(grp == "Population") %>% 
			select_at(sel_abs_names$id) %>% 
			`colnames<-`(sel_abs_names$Name) %>% 
			mutate_all(function(x) if_else(x, 1L, 0L)) %>% 
			as.data.frame() %>% 
			# as.list() %>% 
			# lapply(which) %>% 
			# 
			upset(nsets = 10, 
						sets = rev(paste(c("SPK_02", "RBD_01", "NCP_01"), "-", rep(c("IgG", "IgM"), each = 3))),
						nintersects = 40,
						order.by = "degree",
						keep.order = T,
						text.scale = c(1, 1.5, 1, 1, 1.2, 1),
						# mb.ratio = c(1 - rtio, rtio),
						mainbar.y.label = "Number of seropositive samples [N]") %>% 
			print()
		cat("\n")
	}
}
```


------------------------------------

###  Association with replies to questionnaire


```{r seropos_spk}

seropos_spk <- lapply(names(bBs) %>% `names<-`(., .), function(iset) {
	sel_ab <- bBs[[iset]]$binder %>% 
		filter(detect_Ab == "IgG") %>% 
		filter(grepl("SPK", Name)) %>% 
		pull(id)
	
	posneg[[iset]]$`6` %>% 
		select_at(sel_ab) %>% 
		apply(1, function(x) sum(x) >= 2) %>% 
		enframe(name = NULL, value = "seropos") %>% 
		bind_cols(dplyr::select(posneg[[iset]]$`6`, grp), .) %>% 
		mutate(set = iset) %>% 
		bind_cols(dplyr::select(bBs[[iset]]$sinfo, id), .)
		
}) %>% 
	bind_rows()


seropos_spk %>% 
	group_by(grp, set) %>% 
	summarise(
		n_pos = sum(seropos),
		pos_prop = n_pos / n()
	) %>% 
	dplyr::select(-n_pos) %>% 
	pivot_wider(names_from = grp, values_from = pos_prop) %>% 
	mutate(set = sub("set", "Set", set)) %>% 
	dplyr::select(
		Set = set, 
		"Pos. prop." = Population, 
		Sensitivity = Positive, 
		Specificity = Negative
	) %>% 
	mutate(
		Specificity = 1 - Specificity,
		PPV = (Sensitivity * PREVALENCE) / 
			(Sensitivity * PREVALENCE + (1 - Specificity) * (1 - PREVALENCE))
	) %>% 
	mutate(
		#  P +- dev
		dev = sqrt(`Pos. prop.` * (1 - `Pos. prop.`) / sum(posneg[[iset]][[isd]]$grp == "Population")) * -qnorm(0.025),
		#  lower bound of CI
		lower = `Pos. prop.` - dev,
		lower = if_else(lower < 0, 0, lower),
		#  upper bound of CI
		upper = `Pos. prop.` + dev
	) %>% 
	#  Change to "0.00%" character
	mutate_at(
		c("Pos. prop.", "lower", "upper"),
		function(x) {
			if_else(
				is.na(x),
				"",
				(x * 100) %>% 
					round(., 2) %>% 
					format(width = 5) %>% 
					paste0(., "%")
			)
		}
	) %>% 
	mutate(
		`95% CI` = paste0(lower, "-", upper)
	) %>% 	
	#  Change to "00%" character
	mutate_if(
		is.numeric,
		function(x) {
			if_else(
				is.na(x),
				"",
				(x * 100) %>% 
					round() %>% 
					format(width = 5) %>% 
					paste0(., "%")
			)
		}
	) %>% 
	dplyr::select(
		Set,
		`Pos. prop.`, 
		`95% CI`,
		'Sens' = Sensitivity,
		'Spec' = Specificity,
		PPV
	) %>% 
	kable(caption = paste0("Seropositves")) %>% 
	kable_styling(full_width = F) %>% 
	print()
```

* Wilcoxon test for `Age group` and `Flu-like symptom` to take into account the ordered categorical variables.
* Fisher's exact test for `Sex` and `Breathing difficulty`


```{r qnaire}

tmp <- bBs$set1$sinfo %>% 
	mutate(set = "set1") %>% 
	dplyr::select(id, set, Sex:Breath)
tmp <- bBs$set2$sinfo %>% 
	mutate(set = "set2") %>% 
	dplyr::select(id, set, Sex:Breath) %>% 
	bind_rows(tmp)


tg <- seropos_spk %>% 
	mutate(seropos = if_else(seropos, "Positive", "Negative") %>% 
				 	factor(levels = c("Positive", "Negative", 'P-value'))) %>% 
	left_join(tmp, by= c("id", "set")) %>% 
	filter(grp == "Population") %>% 
	mutate(
		Age_grp = factor(Age_grp, ordered = T, 
										 levels = c("20-29", "30-39", "40-49", "50-59", "60-69", "70-74")),
		Breath = replace(Breath, Breath %in% c("Both", "Coughing", "Hard breat"), "Yes"),
		set = sub("set", "Study set", set)
  )

label(tg$Age_grp) <- "Age group"
units(tg$Age_grp) <- "years"
label(tg$Breath) <- "Breathing difficulty"
label(tg$Symptom) <- "Flu-like symptom"

rndr <- function(x, name, ...) {
  if (length(x) == 0) {
    y <- tg[[name]]
    s <- rep("", length(render.default(x=y, name=name, ...)))
    p <- if(is.ordered(y)) {
    	f <- formula(paste(name, "~ seropos"))
    	tg1 <- tg %>% 
    		mutate(seropos = droplevels(seropos)) %>% 
    		mutate_at(name, as.integer)
    	wilcox.test(f, tg1)$p.value
    } else {
    	fisher.test(table(y, droplevels(tg$seropos), useNA = "no"))$p.value
    }
    
    s[2] <- sub("<", "&lt;", format.pval(p, digits=3, eps=0.001))
    s
  } else {
    render.default(x=x, name=name, ...)
  }
}
rndr.strat <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}

table1(~ Sex + Age_grp + Symptom + Breath | seropos, tg,
       droplevels=F, render= rndr, render.strat=rndr.strat, overall=F)

```

####  Odds ratio

```{r oddsratio}

oddsratio <- function(x) {
	x %>% 
		as.data.frame() %>% 
		pull(Freq) %>% 
		{(.[1] * .[4]) / (.[2] * .[3])} %>% 
		round(2)
}

ortio <- list(
	Sex = tg %>% 
		mutate(seropos = droplevels(seropos)) %>% 
		filter(Sex %in% c("Female", "Male")) %>% 
		{with(., table(seropos, Sex))} %>% 
		oddsratio,
	Symptom = tg %>% 
		mutate(
			seropos = droplevels(seropos),
			Symptom = sub("Yes[, [:alpha:]]*", "Yes", Symptom)
		) %>% 
		{with(., table(Symptom, seropos))} %>% 
		oddsratio,
	Breath = tg %>% 
		mutate(seropos = droplevels(seropos)) %>% 
		{with(., table(seropos, Breath))} %>% 
		oddsratio
)
```

* Sex : `r ortio$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$Breath` (odds of 'No' / odds of 'Yes')


-----------------------------------

###  Seropositive by UMAP

* Wilcoxon test for `Age group` and `Flu-like symptom` to take into account the ordered categorical variables.
* Fisher's exact test for `Sex` and `Breathing difficulty`


```{r umap}
umap <- lapply(
	names(fn$i$umap) %>% `names<-`(., .), 
	function(iset) {
		read.delim(
			fn$i$umap[[iset]],
			sep= " ",
			stringsAsFactors = F
		) %>% 
			as_tibble() %>% 
			mutate(
				set = iset
			) %>% 
			left_join(select(bBs[[iset]]$sinfo, id, grp, Sex:Breath), by = "id")
	}
) %>% 
	bind_rows() %>% 
	mutate(
		seropos = if_else(
			umap_seropositive_IgG_IgM == 1L,
			TRUE,
			FALSE
		)
	) %>% 
	select(-umap_seropositive_IgG_IgM) %>% 
	drop_na(grp)

umap %>% 
	group_by(grp, set) %>% 
	summarise(
		n_pos = sum(seropos),
		pos_prop = n_pos / n()
	) %>% 
	dplyr::select(-n_pos) %>% 
	pivot_wider(names_from = grp, values_from = pos_prop) %>% 
	mutate(set = sub("set", "Set", set)) %>% 
	dplyr::select(
		Set = set, 
		"Pos. prop." = Population, 
		Sensitivity = Positive, 
		Specificity = Negative
	) %>% 
	mutate(
		Specificity = 1 - Specificity,
		PPV = (Sensitivity * PREVALENCE) / 
			(Sensitivity * PREVALENCE + (1 - Specificity) * (1 - PREVALENCE))
	) %>% 
	mutate(
		#  P +- dev
		dev = sqrt(`Pos. prop.` * (1 - `Pos. prop.`) / sum(posneg[[iset]][[isd]]$grp == "Population")) * -qnorm(0.025),
		#  lower bound of CI
		lower = `Pos. prop.` - dev,
		lower = if_else(lower < 0, 0, lower),
		#  upper bound of CI
		upper = `Pos. prop.` + dev
	) %>% 
	#  Change to "0.00%" character
	mutate_at(
		c("Pos. prop.", "lower", "upper"),
		function(x) {
			if_else(
				is.na(x),
				"",
				(x * 100) %>% 
					round(., 2) %>% 
					format(width = 5) %>% 
					paste0(., "%")
			)
		}
	) %>% 
	mutate(
		`95% CI` = paste0(lower, "-", upper)
	) %>% 	
	#  Change to "00%" character
	mutate_if(
		is.numeric,
		function(x) {
			if_else(
				is.na(x),
				"",
				(x * 100) %>% 
					round() %>% 
					format(width = 5) %>% 
					paste0(., "%")
			)
		}
	) %>% 
	dplyr::select(
		Set,
		`Pos. prop.`, 
		`95% CI`,
		'Sens' = Sensitivity,
		'Spec' = Specificity,
		PPV
	) %>% 
	kable(caption = paste0("Seropositves")) %>% 
	kable_styling(full_width = F) %>% 
	print()

tg <- umap %>% 
	mutate(seropos = if_else(seropos, "Positive", "Negative") %>% 
				 	factor(levels = c("Positive", "Negative", 'P-value'))) %>% 
	filter(grp == "Population") %>% 
	mutate(
		Age_grp = factor(Age_grp, ordered = T, 
										 levels = c("20-29", "30-39", "40-49", "50-59", "60-69", "70-74")),
		Breath = replace(Breath, Breath %in% c("Both", "Coughing", "Hard breat"), "Yes"),
		set = sub("set", "Study set", set)
  )

label(tg$Age_grp) <- "Age group"
units(tg$Age_grp) <- "years"
label(tg$Breath) <- "Breathing difficulty"
label(tg$Symptom) <- "Flu-like symptom"

rndr <- function(x, name, ...) {
  if (length(x) == 0) {
    y <- tg[[name]]
    s <- rep("", length(render.default(x=y, name=name, ...)))
    p <- if(is.ordered(y)) {
    	f <- formula(paste(name, "~ seropos"))
    	tg1 <- tg %>% 
    		mutate(seropos = droplevels(seropos)) %>% 
    		mutate_at(name, as.integer)
    	wilcox.test(f, tg1)$p.value
    } else {
    	fisher.test(table(y, droplevels(tg$seropos), useNA = "no"))$p.value
    }
    
    s[2] <- sub("<", "&lt;", format.pval(p, digits=3, eps=0.001))
    s
  } else {
    render.default(x=x, name=name, ...)
  }
}
rndr.strat <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}

table1(~ Sex + Age_grp + Symptom + Breath | seropos, tg,
       droplevels=F, render= rndr, render.strat=rndr.strat, overall=F)



```

####  Odds ratio

```{r oddsratio_umap}

ortio <- list(
	Sex = tg %>% 
		mutate(seropos = droplevels(seropos)) %>% 
		filter(Sex %in% c("Female", "Male")) %>% 
		{with(., table(Sex, seropos))} %>% 
		oddsratio,
	Symptom = tg %>% 
		mutate(
			seropos = droplevels(seropos),
			Symptom = sub("Yes[, [:alpha:]]*", "Yes", Symptom)
		) %>% 
		{with(., table(Symptom, seropos))} %>% 
		oddsratio,
	Breath = tg %>% 
		mutate(seropos = droplevels(seropos)) %>% 
		{with(., table(Breath, seropos))} %>% 
		oddsratio
)
```

* Sex : `r ortio$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$Breath` (odds of 'No' / odds of 'Yes')



-------------------------------------

###  UpSet plots - All SPK,RBD,NCP binders + UMAP


```{r upsetplot_umap, fig.align='center'}


for(isd in names(posneg[[1]])) {
	for(iset in names(posneg)) {
		umap_set <- umap %>% 
			filter(set == iset & grp == "Population") %>% 
			select(subj.id, seropos)
		
		for(i_det_ab in c("IgG", "IgM")) {
			
			cat("\n#### ", sub("set", "Study set ", iset), "-", i_det_ab, " (Cutoff ", isd, " S.D.)\n")

			sel_abs <- (bBs[[iset]]$binder$detect_Ab == i_det_ab) & 
				(grepl("SPK|NCP|RBD", bBs[[iset]]$binder$Name))
			
			sel_abs_names <- bBs[[iset]]$binder %>% 
				dplyr::select(id, Name, detect_Ab) %>% 
				filter(sel_abs) %>% 
				unite(Name, Name, detect_Ab, sep = " - ")
			
			sorted_sel_abs_names <- sel_abs_names %>% 
				separate(Name, c("Name1", "Name2"), sep= "_") %>% 
				arrange(Name1, desc(Name2)) %>% 
				unite(Name, Name1, Name2) %>% 
				pull(Name) %>% 
				c("UMAP", .)

			sel_abs_names <- sel_abs_names %>% 
				add_case(id = "seropos", Name = "UMAP")
			# rtio <- (0.4 - 0.25) / (10 - 4) * (length(sorted_sel_abs_names) - 4) + 0.25
			
			posneg[[iset]][[isd]] %>% 
				filter(grp == "Population") %>% 
				left_join(umap_set, by= "subj.id") %>% 
				select_at(c(sel_abs_names$id, "seropos")) %>% 
				`colnames<-`(sel_abs_names$Name) %>% 
				mutate_all(function(x) if_else(x, 1L, 0L)) %>% 
				as.data.frame() %>% 
				# as.list() %>% 
				# lapply(which) %>% 
				# 
				upset(nsets = 10, 
							sets = sorted_sel_abs_names,
							nintersects = 40,
							order.by = "degree",
							keep.order = T,
							text.scale = c(1, 1.5, 1, 1, 1.2, 1),
							# mb.ratio = c(1 - rtio, rtio),
							mainbar.y.label = "Number of seropositive samples [N]") %>% 
				print()
			cat("\n")
		}
	}
}


```

-------------------------------------

###  UpSet plots - SPK_02, RBD_01, NCP_01 + UMAP


```{r upset_sets_umap, fig.align='center'}


for(isd in names(posneg[[1]])) {
	for(iset in names(posneg)) {
		umap_set <- umap %>% 
			filter(set == iset & grp == "Population") %>% 
			select(subj.id, seropos)
		
		cat("\n#### ", sub("set", "Study set ", iset), " (Cutoff ", isd, " S.D.)\n")
		
		sel_abs <- bBs[[iset]]$binder$Name %in% c("SPK_02", "NCP_01", "RBD_01")
		
		sel_abs_names <- bBs[[iset]]$binder %>% 
			dplyr::select(id, Name, detect_Ab) %>% 
			filter(sel_abs) %>% 
			unite(Name, Name, detect_Ab, sep = " - ") %>% 
			add_case(id = "seropos", Name = "UMAP")
		
		# rtio <- (0.4 - 0.25) / (10 - 4) * (length(sorted_sel_abs_names) - 4) + 0.25
		
		posneg[[iset]][[isd]] %>% 
			filter(grp == "Population") %>% 
			left_join(umap_set, by= "subj.id") %>% 
			select_at(c(sel_abs_names$id, "seropos")) %>% 
			`colnames<-`(sel_abs_names$Name) %>% 
			mutate_all(function(x) if_else(x, 1L, 0L)) %>% 
			as.data.frame() %>% 
			# as.list() %>% 
			# lapply(which) %>% 
			# 
			upset(nsets = 10, 
						sets = c(
							"UMAP",
							rev(paste(c("SPK_02", "RBD_01", "NCP_01"), "-", rep(c("IgG", "IgM"), each = 3)))
						),
						nintersects = 40,
						order.by = "degree",
						keep.order = T,
						text.scale = c(1, 1.5, 1, 1, 1.2, 1),
						# mb.ratio = c(1 - rtio, rtio),
						mainbar.y.label = "Number of seropositive samples [N]") %>% 
			print()
		cat("\n")
	}
}
```




-----------------------------------

###  Seropositive by S-protein/N-protein only

* S-protein only = At least 2 `SPK` positive signals but not positive by UMAP
* N-protein only = At least 1 `NCP` positive but not positive by UMAP

```{r snonly}

spkncp <- lapply(names(posneg[[1]]) %>% `names<-`(., .), function(isd) {
	
	lapply(c("set1", "set2"), function(iset) {
		sel_abs <- (bBs[[iset]]$binder$detect_Ab == "IgG") & 
			(grepl("SPK|NCP", bBs[[iset]]$binder$Name))
		
		sel_abs_names <- bBs[[iset]]$binder %>% 
			dplyr::select(id, Name, detect_Ab) %>% 
			filter(sel_abs) %>% 
			add_case(id = "seropos", Name = "UMAP")
		
		posneg_umap <- posneg[[iset]][[isd]] %>% 
			# filter(grp == "Population") %>% 
			left_join(umap, by= c("sample_id" = "id", "subj.id", "grp")) %>% 
			select(subj.id, grp, one_of(sel_abs_names$id), set:seropos) %>% 
			rename_all(function(x1) {
				i_mat <- match(x1, sel_abs_names$id)
				x1[!is.na(i_mat)] <- sel_abs_names$Name[i_mat[!is.na(i_mat)]]
				x1
			})
		
		posneg_umap %>% 
			mutate(
				#  at least two SPK
				SPK = apply(posneg_umap %>% select(starts_with("SPK")), 1, sum) >= 2 & !UMAP,
				#  at least one NCP
				NCP = apply(posneg_umap %>% select(starts_with("NCP")), 1, sum) >= 1 & !UMAP,
			) %>% 
			select(subj.id, grp, SPK, NCP, set:Breath)
	}) %>% 
		bind_rows()
})	


for(isd in names(spkncp)) {
	cat("\n#### Cutoff ", isd, " S.D.\n")
	
	spkncp[[isd]] %>% 
		pivot_longer(c(SPK, NCP), names_to = "prot", values_to = "seropos") %>% 
		group_by(grp, set, prot) %>% 
		summarise(
			n_pos = sum(seropos),
			pos_prop = n_pos / n()
		) %>% 
		dplyr::select(-n_pos) %>% 
		pivot_wider(names_from = grp, values_from = pos_prop) %>% 
		ungroup() %>% 
		arrange(prot) %>% 
		mutate(set = sub("set", "Set", set)) %>% 
		dplyr::select(
			`Positive by` = prot,
			Set = set, 
			"Pos. prop." = Population, 
			Sensitivity = Positive, 
			Specificity = Negative
		) %>% 
		mutate(
			`Positive by` = paste(`Positive by`, "only"),
			Specificity = 1 - Specificity,
			PPV = (Sensitivity * PREVALENCE) / 
				(Sensitivity * PREVALENCE + (1 - Specificity) * (1 - PREVALENCE))
		) %>% 
		mutate(
			#  P +- dev
			dev = sqrt(`Pos. prop.` * (1 - `Pos. prop.`) / sum(posneg[[iset]][[isd]]$grp == "Population")) * -qnorm(0.025),
			#  lower bound of CI
			lower = `Pos. prop.` - dev,
			lower = if_else(lower < 0, 0, lower),
			#  upper bound of CI
			upper = `Pos. prop.` + dev
		) %>% 
		#  Change to "0.00%" character
		mutate_at(
			c("Pos. prop.", "lower", "upper"),
			function(x) {
				if_else(
					is.na(x),
					"",
					(x * 100) %>% 
						round(., 2) %>% 
						format(width = 5) %>% 
						paste0(., "%")
				)
			}
		) %>% 
		mutate(
			`95% CI` = paste0(lower, "-", upper)
		) %>% 	
		#  Change to "00%" character
		mutate_if(
			is.numeric,
			function(x) {
				if_else(
					is.na(x),
					"",
					(x * 100) %>% 
						round() %>% 
						format(width = 5) %>% 
						paste0(., "%")
				)
			}
		) %>% 
		dplyr::select(
			`Positive by`,
			Set,
			`Pos. prop.`, 
			`95% CI`,
			'Sens' = Sensitivity,
			'Spec' = Specificity,
			PPV
		) %>% 
		kable(caption = paste0("Seropositves")) %>% 
		kable_styling(full_width = F) %>% 
		print()
}

tg <- spkncp %>% 
	lapply(
		. %>% 
			pivot_longer(c(SPK, NCP), names_to = "prot", values_to = "seropos") %>% 
			mutate(seropos = if_else(seropos, "Positive", "Negative") %>% 
						 	factor(levels = c("Positive", "Negative", 'P-value'))) %>% 
			filter(grp == "Population") %>% 
			mutate(
				Age_grp = factor(Age_grp, ordered = T, 
												 levels = c("20-29", "30-39", "40-49", "50-59", "60-69", "70-74")),
				Breath = replace(Breath, Breath %in% c("Both", "Coughing", "Hard breat"), "Yes"),
				set = sub("set", "Study set", set)
			) %>% {
				label(.$Age_grp) <- "Age group"
				units(.$Age_grp) <- "years"
				label(.$Breath) <- "Breathing difficulty"
				label(.$Symptom) <- "Flu-like symptom"
				.
			}
	)

rndr.strat <- function(label, n, ...) {
    ifelse(n==0, label, render.strat.default(label, n, ...))
}
```


* Wilcoxon test for `Age group` and `Flu-like symptom` to take into account the ordered categorical variables.
* Fisher's exact test for `Sex` and `Breathing difficulty`


```{r}
for(pos_by in c("SPK", "NCP")) {
	cat("\n##### ", pos_by, "only\n")
	
	for(isd in names(tg)) {
		cat("\n###### Cutoff ", isd, " S.D.\n")
		
		tg_sd_prot <- tg[[isd]] %>% 
			filter(prot == pos_by)
		rndr <- function(x, name, ...) {
			if (length(x) == 0) {
				y <- tg_sd_prot[[name]]
				s <- rep("", length(render.default(x=y, name=name, ...)))
				p <- if(is.ordered(y)) {
					f <- formula(paste(name, "~ seropos"))
					tg1 <- tg_sd_prot %>% 
						mutate(seropos = droplevels(seropos)) %>% 
						mutate_at(name, as.integer)
					wilcox.test(f, tg1)$p.value
				} else {
					fisher.test(table(y, droplevels(tg_sd_prot$seropos), useNA = "no"))$p.value
				}
				
				s[2] <- sub("<", "&lt;", format.pval(p, digits=3, eps=0.001))
				s
			} else {
				render.default(x=x, name=name, ...)
			}
		}
		
		table1(~ Sex + Age_grp + Symptom + Breath | seropos, tg_sd_prot,
					 droplevels=F, render= rndr, render.strat=rndr.strat, overall=F) %>% 
			print()
	}
}
```


####  Odds ratio

```{r oddsratio_spkncponly}

ortio <- lapply(c("SPK", "NCP") %>% `names<-`(., .), function(pos_by) {
	lapply(names(tg) %>% `names<-`(., .), function(isd) {
	
		tg_sd_prot <- tg[[isd]] %>% 
			filter(prot == pos_by)
		
		list(
			Sex = tg_sd_prot %>% 
				mutate(seropos = droplevels(seropos)) %>% 
				filter(Sex %in% c("Female", "Male")) %>% 
				{with(., table(Sex, seropos))} %>% 
				oddsratio,
			Symptom = tg_sd_prot %>% 
				mutate(
					seropos = droplevels(seropos),
					Symptom = sub("Yes[, [:alpha:]]*", "Yes", Symptom)
				) %>% 
				{with(., table(Symptom, seropos))} %>% 
				oddsratio,
			Breath = tg_sd_prot %>% 
				mutate(seropos = droplevels(seropos)) %>% 
				{with(., table(Breath, seropos))} %>% 
				oddsratio
		)
	})
})
```

#####  SPK only

######  Cutoff 3 S.D.

* Sex : `r ortio$SPK[['3']]$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$SPK[['3']]$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$SPK[['3']]$Breath` (odds of 'No' / odds of 'Yes')

######  Cutoff 6 S.D.

* Sex : `r ortio$SPK[['6']]$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$SPK[['6']]$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$SPK[['6']]$Breath` (odds of 'No' / odds of 'Yes')


#####  NCP only

######  Cutoff 3 S.D.

* Sex : `r ortio$NCP[['3']]$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$NCP[['3']]$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$NCP[['3']]$Breath` (odds of 'No' / odds of 'Yes')

######  Cutoff 6 S.D.

* Sex : `r ortio$NCP[['6']]$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$NCP[['6']]$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$NCP[['6']]$Breath` (odds of 'No' / odds of 'Yes')


-------------------------------------

###  UpSet plots - SPK,NCP + UMAP


```{r upset_sets_spkncponly, fig.align='center'}


for(isd in names(spkncp)) {
	cat("\n#### Cutoff ", isd, " S.D.\n")
	
	spkncp[[isd]] %>% 
		filter(grp == "Population") %>% 
		left_join(umap %>% filter(grp == "Population"), by= "subj.id") %>% 
		select("SPK only" = "SPK", "NCP only" = "NCP", UMAP = "seropos") %>% 
		mutate_all(function(x) if_else(x, 1L, 0L)) %>% 
		as.data.frame() %>% 
		# as.list() %>% 
		# lapply(which) %>% 
		# 
		upset(nsets = 10, 
					sets = c("UMAP", "NCP only", "SPK only"),
					nintersects = 40,
					order.by = "degree",
					keep.order = T,
					text.scale = c(1, 1.5, 1, 1, 1.2, 1),
					# mb.ratio = c(1 - rtio, rtio),
					mainbar.y.label = "Number of seropositive samples [N]") %>% 
		print()
	cat("\n")
}
```


-------------------------------------

###  Comparison of S-protein/N-protein only and UMAP

* The Cutoff of 6 S.D. was chosen here.

* Wilcoxon test for `Age group` and `Flu-like symptom` to take into account the ordered categorical variables.
* Fisher's exact test for `Sex` and `Breathing difficulty`


```{r snumap}

umap1 <- umap %>% 
	filter(grp == "Population") %>% 
	select(subj.id, UMAP = seropos)

for(pos_by in c("SPK", "NCP")) {
	cat("\n##### ", pos_by, "only\n")
	
	tg_sd_prot <- tg[['6']] %>% 
		filter(grp == "Population") %>% 
		left_join(umap1, by= "subj.id") %>% 
		filter(prot == pos_by) %>% 
		mutate(
			seropos = case_when(
				seropos == "Positive" ~ paste(pos_by, "only"),
				UMAP ~ "UMAP",
				TRUE ~ NA_character_
			) %>% 
				factor(levels = c(paste(pos_by, "only"), "UMAP", 'P-value'))
		) %>% 
		filter(!is.na(seropos))
	

	rndr <- function(x, name, ...) {
		if (length(x) == 0) {
			y <- tg_sd_prot[[name]]
			s <- rep("", length(render.default(x=y, name=name, ...)))
			p <- if(is.ordered(y)) {
				f <- formula(paste(name, "~ seropos"))
				tg1 <- tg_sd_prot %>% 
					mutate(seropos = droplevels(seropos)) %>% 
					mutate_at(name, as.integer)
				wilcox.test(f, tg1, exact = F)$p.value
			} else {
				fisher.test(table(y, droplevels(tg_sd_prot$seropos), useNA = "no"))$p.value
			}
			
			s[2] <- sub("<", "&lt;", format.pval(p, digits=3, eps=0.001))
			s
		} else {
			render.default(x=x, name=name, ...)
		}
	}
	
	table1(~ Sex + Age_grp + Symptom + Breath | seropos, tg_sd_prot,
				 droplevels=F, render= rndr, render.strat=rndr.strat, overall=F) %>% 
		print()
}


```


####  Odds ratio

```{r oddsratio_snumap}

ortio <- lapply(c("SPK", "NCP") %>% `names<-`(., .), function(pos_by) {
	lapply(names(tg) %>% `names<-`(., .), function(isd) {
	
		tg_sd_prot <- tg[[isd]] %>% 
			filter(grp == "Population") %>% 
			left_join(umap1, by= "subj.id") %>% 
			filter(prot == pos_by) %>% 
			mutate(
				seropos = case_when(
					seropos == "Positive" ~ paste(pos_by, "only"),
					UMAP ~ "UMAP",
					TRUE ~ NA_character_
				) %>% 
					factor()
			) %>% 
			filter(!is.na(seropos))

		list(
			Sex = tg_sd_prot %>% 
				mutate(seropos = droplevels(seropos)) %>% 
				filter(Sex %in% c("Female", "Male")) %>% 
				{with(., table(Sex, seropos))} %>% 
				oddsratio,
			Symptom = tg_sd_prot %>% 
				mutate(
					seropos = droplevels(seropos),
					Symptom = sub("Yes[, [:alpha:]]*", "Yes", Symptom)
				) %>% 
				{with(., table(Symptom, seropos))} %>% 
				oddsratio,
			Breath = tg_sd_prot %>% 
				mutate(seropos = droplevels(seropos)) %>% 
				{with(., table(Breath, seropos))} %>% 
				oddsratio
		)
	})
})
```

#####  SPK only

* Sex : `r ortio$SPK[['6']]$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$SPK[['6']]$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$SPK[['6']]$Breath` (odds of 'No' / odds of 'Yes')



#####  NCP only

######  Cutoff 6 S.D.

* Sex : `r ortio$NCP[['6']]$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$NCP[['6']]$Symptom`  (odds of 'No' / odds of 'Yes')

* Breath : `r ortio$NCP[['6']]$Breath` (odds of 'No' / odds of 'Yes')



```{r sessionInfo, results='markup'}
sessionInfo()
```




