---
title: "anal-cbc03-seropos_prop"
author: "Mun-Gwan"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  html_document:
    theme: readable
    toc: true
    toc_float: true
    toc_depth: 3
    fig_caption: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../results/sero_positive_by_sd") })
---

<style>
p.caption {
  font-size: 1em;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	dev = c('png', 'pdf'),
	fig.path = '../results/sero_positive_by_sd/figures/',  # save figure in files
	results = 'asis'
)

library(tidyverse)
library(Useful2me)  # remotes::install_github("Rundmus/Useful2me-R_package", upgrade= "never")
library(knitr)
library(kableExtra)     # kable_styling
library(UpSetR)
library(table1)

# library(janitor)     # tabyl, fisher.test

rm(list = ls())


###  File names

source("options_master_trans_serology.R")
# setwd("~/covid19_serology/scripts")

fn <- list(
	i = list(                               #  input
		cbc3 = data_file_path_cbc03,
		binder = '../data/assay_information/bead_id_information/CBA5_information_js_matching.csv',
		umap = '../results/dim_reduction_plots/cbc03/luminex/2020-07-17_dim_red_dbs_covid19_manuscript_cbc03_cba_v1_detectionAb_IgG_seropositivity.txt'
	)
)

#  Check all exists
stopifnot(all(file.exists(c('.', unlist(fn$i)))))


###  Load data

tmp <- load(fn$i$cbc3)
bBs0 <- list(
	sinfo = sinfo,
	mfi = mfi,
	binder = binder
)
rm(list = tmp)

bBs <- bBs0


cbas <- read_csv(
	fn$i$binder,
	col_types = cols_only(
		# CBA_nr = col_character(),
		# Binder = col_character(),
		# Comment = col_character(),
		Bead_ID = col_character(),
		# Coupling_chemistry = col_character(),
		Name = col_character(),
		# Provider = col_character(),
		# `Edsberg Analysis` = col_character(),
		`In batch 1` = col_character(),
		`Batch 2 Analysis` = col_character(),
		Match_Batch_1 = col_character()
	)
)
stopifnot(anyDuplicated(cbas$Name) == 0)


umap0 <- read.delim(
	fn$i$umap,
	sep= " ",
	stringsAsFactors = F
) %>% 
	as_tibble()
umap <- umap0 %>% 
	mutate(
		UMAP = if_else(
			umap_seropositive_IgG == 1L,
			TRUE,
			FALSE
		)
	) %>% 
	select(-umap_seropositive_IgG)

##  CONSTANTS

SERONEG_PROP = 0.8    # assumed sero-negative proportion 
PREVALENCE = 0.1      # assumed the prevalence of Covid-19
SD_CUTOFF = 6
```

```{r sinfo_umap}

stopifnot(anyDuplicated(umap$id) == 0)

bBs$sinfo <- bBs0$sinfo %>% 
	left_join(umap, by= c("id", "subj.id")) %>% 
	mutate(
		subj.id_subtype = paste(subj.id, subtype),
		subtype = if_else(duplicated(subj.id_subtype), paste0(subtype, "2"), subtype)
	) %>% 
	select(-subj.id_subtype)
	
stopifnot(all(!is.na(bBs$sinfo$UMAP)))
```

###  UMAP seropositivity


```{r umap_seropos}

bBs$sinfo %>% 
	group_by(subtype) %>% 
	summarise(
		n = n(),
		n_pos = sum(UMAP),
		pos_prop = n_pos / n()
	) %>% 
	dplyr::select(-n_pos) %>% 
	# pivot_wider(names_from = subtype, values_from = pos_prop)# %>% 
	dplyr::select(
		n,
		Subtype = subtype, 
		"Pos. prop." = pos_prop 
	) %>% 
	mutate(
		#  P +- dev
		dev = sqrt(`Pos. prop.` * (1 - `Pos. prop.`) / n) * -qnorm(0.025),
		#  lower bound of CI
		lower = `Pos. prop.` - dev,
		lower = if_else(lower < 0, 0, lower),
		#  upper bound of CI
		upper = `Pos. prop.` + dev
	) %>%
	dplyr::select(-n) %>% 
	#  Change to "0.00%" character
	mutate_at(
		c("Pos. prop.", "lower", "upper"),
		function(x) {
			if_else(
				is.na(x),
				"",
				(x * 100) %>% 
					round(., 2) %>% 
					format(width = 5) %>% 
					paste0(., "%")
			)
		}
	) %>% 
	mutate(
		`95% CI` = paste0(lower, "-", upper)
	) %>% 	
	#  Change to "00%" character
	mutate_if(
		is.numeric,
		function(x) {
			if_else(
				is.na(x),
				"",
				(x * 100) %>% 
					round() %>% 
					format(width = 5) %>% 
					paste0(., "%")
			)
		}
	) %>% 
	dplyr::select(
		Subtype,
		`Pos. prop.`, 
		`95% CI`
	) %>% 
	kable(caption = paste0("Seropositves")) %>% 
	kable_styling(full_width = F) %>% 
	print()
```

*  `Plasma` and `cDBS` are used. All other sample types were neglected.

```{r}
bBs <- lapply(
	list("cDBS", "Plasma") %>% `names<-`(., .),
	function(ii) {
		list(
			sinfo = bBs$sinfo %>% 
				filter(subtype == ii),
			mfi = bBs$mfi[bBs$sinfo$subtype == ii, ],
			binder = bBs$binder
		)
	}
)
```


##  6 Std.Dev from the peak (like CBC-01)

```{r}

##  Sero-positive or not (T or F) matrix
posneg <- lapply(names(bBs) %>% `names<-`(., .), function(iset) {
	#  No duplicated CBC+03 samples
	stopifnot(
		anyDuplicated(
			bBs[[iset]]$sinfo$subj.id[grepl("CBC\\+", bBs[[iset]]$sinfo$subj.id)]
		) == 0
	)

	subj.id <- bBs[[iset]]$sinfo$subj.id
	id <- bBs[[iset]]$sinfo$id
	UMAP <- bBs[[iset]]$sinfo$UMAP
	
	#  Finding cutoff ---------------------------------------------------------

	mfi_pop <- bBs[[iset]]$mfi %>% 
		as_tibble()

	##  Based on peak value
	den <- mfi_pop %>%
		lapply(density)

	#  peak
	m <- sapply(den, function(x) x$x[which.max(x$y)])

	#  std.dev. of (SERONEG_PROP) around the peak
	s <- sapply(seq(m), function(ii) {
		df <- abs(mfi_pop[[ii]] - m[ii])
		qtl <- quantile(df, probs = c(SERONEG_PROP))
		sd(mfi_pop[[ii]][df <= qtl])
	})


	#  Cutoff = peak + 6 * std.dev. of (SERONEG_PROP) around the peak
	cutoff <- m + 6 * s
	
	#  Positve = TRUE, Negative = FALSE
	truefalse <- function(coff, mat) {
		rep(coff, each= nrow(mat)) %>%
			matrix(., nrow= nrow(mat)) %>%
			{mat > .} %>%
			as_tibble()
	}
	
	truefalse(cutoff, bBs[[iset]]$mfi) %>%
		mutate(sample_id = id, subj.id = subj.id, UMAP = UMAP) %>%
		dplyr::select(sample_id, subj.id, UMAP, everything())
})
```

###  Tables

```{r table}
for(iset in names(posneg)) {
	
	out <- posneg[[iset]] %>% 
		
		pivot_longer(-c(sample_id, subj.id, UMAP), names_to = "id") %>%
		group_by(id) %>%
		summarise(
			Sero_pos = sum(value),        # sero-positive
			Non_reac = sum(!value)        # non-reactive (= sero-negative)
		) %>%
		mutate(
			#  Add binder info
			i_match = match(id, bBs[[iset]]$binder$id),
			Binder = bBs[[iset]]$binder$Binder[i_match],
			Name = bBs[[iset]]$binder$Name[i_match],
			det_ab = bBs[[iset]]$binder$detect_Ab[i_match],
			#  Positive proportion
			pos_prop = Sero_pos / (Sero_pos + Non_reac)
		) %>%
		dplyr::select(-i_match) %>% 
		arrange(det_ab) %>% 
		ungroup() 
	
	out1 <- out %>% 
		select(id, pos_prop) %>% 
		rename("Pos. prop." = pos_prop) %>% 
		mutate(
			#  P +- dev
			dev = sqrt(`Pos. prop.` * (1 - `Pos. prop.`) / nrow(posneg[[iset]])) * -qnorm(0.025),
			#  lower bound of CI
			lower = `Pos. prop.` - dev,
			lower = if_else(lower < 0, 0, lower),
			#  upper bound of CI
			upper = `Pos. prop.` + dev
		) %>% 
		#  Change to "0.00%" character
		mutate_at(
			c("Pos. prop.", "lower", "upper"),
			function(x) {
				if_else(
					is.na(x),
					"",
					(x * 100) %>% 
						round(., 2) %>% 
						format(width = 5) %>% 
						paste0(., "%")
				)
			}
		) %>% 
		mutate(
			`95% CI` = paste0(lower, "-", upper)
		) %>% 	
		#  Change to "00%" character
		mutate_if(
			is.numeric,
			function(x) {
				if_else(
					is.na(x),
					"",
					(x * 100) %>% 
						round() %>% 
						format(width = 5) %>% 
						paste0(., "%")
				)
			}
		)
	
	out %>% 
		left_join(out1, by= "id") %>% 
		dplyr::select(
			"Detect Ab" = det_ab, 
			Binder, 
			Name,
			`Pos. prop.`, 
			`95% CI`,
			'Pos.' = Sero_pos, 
			'Neg.' = Non_reac
		) %>% 
		kable(caption = paste0("Seropositves in ", iset, " (Cutoff 6 S.D.)")) %>% 
		kable_styling(full_width = F) %>% 
		print()
}
```

-----------------------------------

###  UMAP

* Wilcoxon test for `Age group` and `Flu-like symptom` to take into account the ordered categorical variables.
* Fisher's exact test for `Sex` and `Breathing difficulty`

```{r qnaire}
for(iset in names(posneg)) {
	cat("\n\n#####", iset, "\n\n")

	# tg <- seropos_spk %>% 
	# 	mutate(seropos = if_else(seropos, "Positive", "Negative") %>% 
	# 				 	factor(levels = c("Positive", "Negative", 'P-value'))) %>% 
	# 	left_join(tmp, by= c("id", "set")) %>% 
	# 	filter(grp == "Population")
	tg <- bBs[[iset]]$sinfo %>% 
		mutate(
			Age_grp = factor(Age_grp, ordered = T, 
											 levels = c("20-29", "30-39", "40-49", "50-59", "60-69", "70-74")),
			Breath = replace(Breath, Breath %in% c("Both", "Coughing", "Hard breat"), "Yes"),
			UMAP = if_else(UMAP, "Positive", "Negative") %>% 
				factor(levels = c("Positive", "Negative", 'P-value'))
		)

	label(tg$Age_grp) <- "Age group"
	units(tg$Age_grp) <- "years"
	label(tg$Breath) <- "Breathing difficulty"
	label(tg$Symptom) <- "Flu-like symptom"
	
	rndr <- function(x, name, ...) {
		if (length(x) == 0) {
			y <- tg[[name]]
			s <- rep("", length(render.default(x=y, name=name, ...)))
			p <- if(is.ordered(y)) {
				f <- formula(paste(name, "~ UMAP"))
				tg1 <- tg %>% 
					mutate(UMAP = droplevels(UMAP)) %>% 
					mutate_at(name, as.integer)
				wilcox.test(f, tg1, exact = F)$p.value
			} else {
				fisher.test(table(y, droplevels(tg$UMAP), useNA = "no"))$p.value
			}
			
			s[2] <- sub("<", "&lt;", format.pval(p, digits=3, eps=0.001))
			s
		} else {
			render.default(x=x, name=name, ...)
		}
	}
	rndr.strat <- function(label, n, ...) {
		ifelse(n==0, label, render.strat.default(label, n, ...))
	}
	
	print(
		table1(~ Sex + Age_grp + Symptom + Breath | UMAP, tg,
					 droplevels=F, render= rndr, render.strat=rndr.strat, overall=F)
	)
}
```

####  Odds ratio

```{r oddsratio}

oddsratio <- function(x) {
	x %>% 
		as.data.frame() %>% 
		pull(Freq) %>% 
		{(.[1] * .[4]) / (.[2] * .[3])}
}

ortio <- lapply(names(bBs) %>% `names<-`(., .), function(iset) {
	tg_sd_prot <- bBs[[iset]]$sinfo %>% 
		mutate(UMAP = if_else(UMAP, "Positive", "Negative") %>% 
				factor(levels = c("Positive", "Negative", 'P-value')))
	
	list(
		Sex = tg_sd_prot %>% 
			mutate(UMAP = droplevels(UMAP)) %>% 
			filter(Sex %in% c("Female", "Male")) %>% 
			{with(., table(Sex, UMAP))} %>% 
			oddsratio %>% 
			round(2),
		Symptom = tg_sd_prot %>% 
			mutate(
				UMAP = droplevels(UMAP),
				Symptom = sub("Yes[, [:alpha:]]*", "Yes", Symptom)
			) %>% 
			{with(., table(Symptom, UMAP))} %>% 
			oddsratio %>% 
			{1/.} %>% 
			round(2),
		Breath = tg_sd_prot %>% 
			mutate(UMAP = droplevels(UMAP)) %>% 
			{with(., table(Breath, UMAP))} %>% 
			oddsratio %>% 
			round(2)
	)
})
```

#####  cDBS

* Sex : `r ortio$cDBS$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$cDBS$Symptom`  (odds of 'Yes' / odds of 'No')

* Breath : `r ortio$cDBS$Breath` (odds of 'No' / odds of 'Yes')


#####  Plasma

* Sex : `r ortio$Plasma$Sex`  (odds for seropositive of 'Female' / odds of 'Male')

* Symptom : `r ortio$Plasma$Symptom`  (odds of 'Yes' / odds of 'No')

* Breath : `r ortio$Plasma$Breath` (odds of 'No' / odds of 'Yes')





```{r sessionInfo, results='markup'}
sessionInfo()
```




