---
title: "Beeswarm plots per antigen"
author: "Creator: Tea Dodig-Crnkovic. Contributors: Mun-Gwan Hong, Cecilia Engel Thomas, Tea Dodig-Crnkovic"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false 

---
<!-- highlight: textmate -->

<!-- Edit TOC style (background color) -->
<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #EAEAEA;
    color: black;}
</style>

<!-- Set background color for highlighted text -->
<style>
div.highlight { 
  background-color: #EAEAEA;  <!-- #39AC91 (turquoise) -->
  color: #FFFFEF;
  padding: 10px;
  border-radius: 2px 2px 2px 2px;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE)
```


<style type="text/css">

h1.title {
  font-size: 22px;
}
h1 { /* Header 1 */
  font-size: 20px;
}
h2 { /* Header 2 */
    font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 12px;
}

</style>

  
# Initialization
```{r eval=T, warning=F, message=FALSE}

# Empty global environment
rm(list = ls(all.names = TRUE))

```

## Load global options
```{r eval=T, warning=F, message=FALSE}

# set global options
source("options_master_trans_serology.R")

#if there was a global data set defined, use that 
if(global_data_set != "no_global_data_set_defined") { 
  data_set = global_data_set 
  
  #otherwise use a localy defined data set, this option is primarily used for individual testing of the analysis scripts  
} else { 
  data_set = "batch2" 
} 

# get path for chosen data set
data_file_path = get_data_set_path(data_set)[[1]]
exp_method = get_data_set_path(data_set)[[2]]

# load data
load(data_file_path)

# output directory
out_dir <- paste0("../results/beeswarms_per_ag/", data_set, "/", exp_method, "/")

# set plot day (today)
plot_date <- format(Sys.time(),"%Y-%m-%d")


```

```{r eval=T, warning=F, message=FALSE}
# Load packages
library(tidyverse)
library(ggbeeswarm)
library(reshape2)
library(ggforce)
library(scales)

options(stringsAsFactors = FALSE)

# exclude bindeders according to list from 2020-06-10
excl <- c(grep("PEP", binder$Name),
          grep("GPE", binder$Name),
          grep("IGA", binder$Name),
          grep("ACE",  binder$Name),
          grep("BBD",  binder$Name),
          grep("NTR",  binder$Name))

if(data_set == "batch1"){
  
  excl <- c(excl,
            c(grep("IGM_03", binder$Name),
              grep("IGM_04", binder$Name),
              grep("PRS", binder$Name),
              grep("NBT", binder$Name)))
  
}

binder <- binder[-excl, ]
mfi <- mfi[, intersect(colnames(mfi), binder$id)]

# Make uniqe Name + detection ab
binder$Name_detect_Ab <- paste0(binder$Name, "_", binder$detect_Ab)

# Melt mfi to long format and 
# create column with unique identifier (sample.antigen)
df <- melt(mfi, value.name = "mfi") %>% 
  unite("unique.id", Var1:Var2, sep=".", remove = FALSE)


# Rename columns
df <-
  df %>% rename("sample.id" = "Var1",
                "protein.id" = "Var2")

sinfo <-
  sinfo %>% rename("sample.id" = "id")

binder <-
  binder %>% rename("protein.id" = "id")


# Merge with protein and sample information
df$protein.id <- as.character(df$protein.id)
df$sample.id <- as.character(df$sample.id)

df <-
  left_join(df, binder, by="protein.id")

df <-
  left_join(df, sinfo, by="sample.id")

# Exclude samples that
#   *) Update 2020-04-29: only include DBS samples 
#   *) Blood spot not OK

df <- df[df$sample.id %in% sinfo$sample.id[sinfo$sample_type == "DBS"],]

exl <- c(sinfo$sample.id[sinfo$sample_type == "DBS" & sinfo$Nr_OK_spots == 0]) %>% unique()

df <- df[!df$sample.id %in% exl, ]



```


# Generate MFI plots
```{r results = "hide"}

# Color by sample type

# split plottong by detection antibody
for(det_Ab in unique(binder$detect_Ab)){
  
pdf(paste0(out_dir,
           format(Sys.time(),"%Y-%m-%d_%H%M%S"),
           "_", data_set, "_beeswarm_panel_mfi_", det_Ab, ".pdf"),
    width=12,height=8, useDingbats=F)
  
  plot_list <- list(1:length(unique(binder$protein.id[binder$detect_Ab == det_Ab])))

# Specify how many plot to do per page
ncol_n <- 4
nrow_n <- 3

n_pages_needed <-
  ceiling(length(unique(df$Name_detect_Ab[df$detect_Ab == det_Ab]))/(ncol_n*nrow_n))


for (i in seq_len(n_pages_needed)) {

  # select protein name and detection Ab
  protein <- unique(df$Name_detect_Ab[df$detect_Ab == det_Ab])
  df_plot <- df[df$Name_detect_Ab %in% protein,]
  

  # Exclude NaN
  df_plot <- df_plot[complete.cases(df_plot$mfi), ]

  # Save panel of plots to list per page
  plot_list[[i]] <-
    ggplot(transform(df_plot, 
                     Name_detect_Ab = factor(Name_detect_Ab, levels= sort(protein))), #sort by alphabetical name
           aes(x = 0, y = mfi, color= sample_type)) +
    geom_beeswarm(cex=1.5, alpha=0.4, priority = "none") +
    xlab("") +
    ylab("MFI[AU]") +
    theme_bw() +
    #scale_y_continuous(trans = 'log2, breaks = log_breaks())
    theme(axis.ticks.x = element_blank(),
          axis.text.x = element_blank()) +
    labs(caption = paste0("Plot date: ", plot_date,
                          "\nInput data: ", data_set)) +
    facet_wrap_paginate(~Name_detect_Ab, 
                        ncol = ncol_n, nrow = nrow_n, page = i,
                        strip.position="top", scales="free_y") +
     scale_color_manual(values=c("black"))
  
 print(plot_list[[i]])
  
}
  
  dev.off()
}


```



# Session information
```{r session_info, eval=T, echo=F}
sessionInfo()

```
