---
title: "Dimensionality reduction of Luminex data"
author: "Creator: Cecilia Engel Thomas. Contributors: Mun-Gwan Hong, Tea Dodig-Crnkovic, Cecilia Engel Thomas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    code_folding: hide
---



# Initialization

```{r packages, eval=t, echo=F}

library(ggplot2)
library(DT)
library(umap)
library(stringr)
library(tidyverse)
library(cowplot)
library(ggpubr)
library(patchwork)

```

```{r initialization, eval=T, warning=F}

rm(list=ls())

options(stringsAsFactors = FALSE)


```

## Load global options

<details><summary>Click here to see options</summary>
```{r engine='bash', comment=''}

#display code from global options
cat options_master_trans_serology.R
```
</details>


```{r global_options, eval=T}

#also sourcing as otherwise the script doesn't work when running the code from R studio as opposed to rendering the report. For rendering this is not neccessary.
source("options_master_trans_serology.R")

#--- Define which data set to run analysis on

#if there was a global data set defined, use that
if(global_data_set != "no_global_data_set_defined") {
  data_set = global_data_set
  
#otherwise use a localy defined data set, this option is primarily used for individual testing of the analysis scripts 
} else {
  data_set = "cbc03"
}

#--- Define which detection antibodies to include in analysis

#if there was a setting, use that
if(global_data_set != "no_global_detection_ab_defined") {
  detection_ab = global_detection_ab
  
#otherwise use a localy setting, this option is primarily used for individual testing of the analysis scripts 
} else {
  detection_ab = "IgG_IgM"
}

#--- Get path for chosen data set
data_file_path = get_data_set_path(data_set)[[1]]
exp_method = get_data_set_path(data_set)[[2]]

#--- Define output directory
out_dir = paste0("../results/dim_reduction_plots/", data_set, "/", exp_method, "/")

```

## Local options

If data is cbc03 then:   
- Data is log2 transformed.   
- All samples except buffer are included (as there are also plasma samples etc in this data set which we want to include).   
- Antibodies targeting CoV proteins (not peptides) are included as well as controls.   

If data is batch1 or batch2 then:   
- Data is not transformed as it is assumed that this has already been done.   
- Only samples that have been on DBS are included.   
- Only antibodies targeting CoV related proteins (not peptides) are included.   


```{r local_options}

# Options
if(data_set == "cbc03"){
  transformation = "log2"
  samples_to_include = "All except buffer"   #"Only DBS from random sampling", "All samples that have been on DBS", "All samples", "All except buffer"
  binders_to_include = "CoV, controls, no peptides"  #"all", "CoV, no peptides", "CoV, controls, no peptides"
} else if (data_set %in% c("batch1","batch2")){
  transformation = "none"
  samples_to_include = "All samples that have been on DBS" #"Only DBS from random sampling", "All samples that have been on DBS", "All samples", "All except buffer
  binders_to_include = "CoV, no peptides"  #"all", "CoV, no peptides", "CoV, controls, no peptides"
}
#samples_to_include = "All except buffer"   #"Only DBS from random sampling", "All samples that have been on DBS", "All samples", "All except buffer"
#binders_to_include = "CoV, controls, no peptides"  #"all", "CoV, no peptides", "CoV, controls, no peptides"

# Plot file prefix
data_set_name = data_file_path %>% str_split(., "/") %>% unlist() %>% .[length(.)] %>% str_split(., "\\.") %>% unlist() %>% .[1]
plot_prefix = paste0(out_dir, Sys.Date(), "_dim_red_", data_set_name, "_detectionAb_", detection_ab)

# Plot footnote
footnote = paste(samples_to_include, paste0("Detection Ab: ", detection_ab), paste0("Included binders: ", binders_to_include), paste0("Transformation: ", transformation), sep=". ")

```



**Working directory**:   
*`r getwd()`*   

**All plots generated can be found at**:   
*`r normalizePath("../results/dim_reduction_plots/")`*  

**Data options**

*Data set*:   
`r data_set`

*Data set input file*:   
`r data_file_path`

*Detection antibodies included in analysis*:   
`r detection_ab`

*Binders included in analysis:*   
`r binders_to_include`

*Samples included in analysis:*   
`r samples_to_include`

*Data transformations applied:*   
`r transformation`

# Read in data

```{r read_data, eval=T}

# MFI, probe info and sample info
load(data_file_path)

#objects: binder, median_bead_count, mfi, min_bead_count, sinfo

mfi_input = mfi

```

# Remove unwanted samples

Exclude samples that   
* Blood spot not OK    
* Only include samples as defined in options (see above)

```{r remove_samples, eval=T}

exclude = c(sinfo$id[sinfo$sample_type == "DBS" & sinfo$Nr_OK_spots == 0]            #DBS samples with no ok blood spot
            )

#if only "random" DBS samples should be included
if(samples_to_include == "Only DBS from random sampling") {
  exclude = c(exclude, sinfo$id[!(sinfo$sample_type %in% c("DBS"))])   #Non DBS saples
}

#if only DBS samples should be included
if(samples_to_include == "All samples that have been on DBS") {
  exclude = c(exclude, sinfo$id[!(sinfo$sample_type %in% c("DBS","DBS from NR", "Serum_DBS", "xDBS"))])   #Non DBS saples
}

# If all samples but buffer should be included
if(samples_to_include == "All except buffer") {
  exclude = sinfo$id[sinfo$sample_type %in% c("Buffer")]  #Buffer
}

# If data is cbc03, remove replicated samples
if(data_set == "cbc03"){
  
  exclude = c(exclude, sinfo$id %>% grep("\\*", ., value=T))   #Replicated samples
  exclude = c(exclude, sinfo$id[sinfo$subtype == "xDBS"])   #xDBS saples
  
}

samples_to_keep = mfi %>%
                      rownames() %>%
                      .[!(. %in% exclude)]

```

**Samples included in analysis**   
`r sinfo %>%
    filter(id %in% samples_to_keep) %>%
    .[,c("id","class","sample_type","Nr_OK_spots")] %>%
    datatable()`


# Filter binders

```{r filter_binders, echo=F}

#--- Filter on detection antibody

if(detection_ab != "all") {
  
  #identify which detection antibodies to include
  abs = detection_ab %>% str_split(., "_") %>% unlist()
  
  mfi_input = mfi_input[,
                binder %>%
                  filter(detect_Ab %in% abs) %>%
                  pull(id)
                ]
  
}

#fix names of binders (ggplot doesn't like the * and parenthesis characters)
colnames(mfi_input) = mfi_input %>% colnames() %>% make.names()
binder = binder %>%
          mutate(id = make.names(id))

#--- Filter on binder identity (control, anti-SARS etc.)

binders_to_keep = mfi_input %>% 
                  colnames()

if(binders_to_include != "all") {

  if(binders_to_include == "CoV, no peptides" | binders_to_include == "CoV, controls, no peptides"){
    
    binders_to_keep = mfi_input %>% 
                  colnames() %>%
                  grep("CoV", ., value=T) %>%              #Only CoV proteins
                  grep("Peptide", ., value=T, invert=T)    #No peptides
      
  } 
  if(binders_to_include == "CoV, controls, no peptides"){
    
    binders_to_keep = c(binders_to_keep, 
            mfi_input %>%
               colnames() %>%
               grep("anti.human", ., value = T),         #Positive controls
            mfi_input %>%
               colnames() %>%
               grep("Bare.Bead", ., value=T)             #Negative controls
            )
    
  }
  
}

```

**Binders included in dimensionality reduction computation**   

`r binder %>%
    filter(id %in% binders_to_keep) %>%
    datatable()`


# Data transformations

```{r data_transform}

if(transformation == "log2"){
  
  mfi_input = mfi_input %>%
                log2(.)
  
}

```

# UMAP function, hyperparameter exploration

Hyperparamaters explored are: 
n_neighbors: "controls how UMAP balances local versus global structure in the data"   
min_dist: "controls how tightly UMAP is allowed to pack points together"   

Euclidean distance used as distance metric.

Additionally, 3 repeats are performed for each combination of hyperparameters. Each repeat has a unique, defined seed.

Further information on UMAP hyperparamaters can be found here: "https://umap-learn.readthedocs.io/en/latest/parameters.html"

```{r umap_function}

umap_hyperparam_explo <- function(data, n_neighbors, min_dist, seeds){
  
  #list for storing results
  results = list()
  
  #looping over n_neighbors
  for(n_neigh in n_neighbors) {
    
    #looping over min_dist
    for(m_dist in min_dist) {
      
      #lopping over seeds
      for(s in seeds) {
        
        name = paste0("n_neighbors: ", n_neigh, ", min_dist: ", m_dist, ", seed: ", s)
        
        results[[name]] = umap(dim_red_input, n_neighbors = n_neigh, min_dist = m_dist, random_state=s)$layout %>% as.data.frame
        
      }
    }
  }
  
  return(results)
  
  
}

```


# Compute dimensionality reductions

```{r compute_dimred, echo=F, warning=F}

#list for storing dimensionality reductions
dim_reds = list()

#data input for dimensionality reduction
dim_red_input = mfi_input[samples_to_keep, binders_to_keep]

#--- PCA

#samples
dim_reds[["samples"]][["pca"]][["1"]] = prcomp(na.omit(dim_red_input))$x[,1:2]

#probes
dim_reds[["probes"]][["pca"]][["1"]] = prcomp(t(na.omit(dim_red_input)))$x[,1:2]

#--- UMAP

#Number of neighbors determined based on number of observations. 
#This is done since the number of neighbors must be less than the number of observations.
#Since the number of neighbors must be >1, 1 is added to comply with this.

#samples
n_neighbors = c(2, ceiling(nrow(dim_red_input)*(1/20))+1, ceiling(nrow(dim_red_input)*(1/3))+1)
dim_reds[["samples"]][["umap"]] = umap_hyperparam_explo(dim_red_input, n_neighbors=n_neighbors, min_dist=c(0.01,0.1,0.5), seeds=c(42,59,254))

#probes
n_neighbors = c(2, ceiling(ncol(dim_red_input)*(1/20))+1, ceiling(ncol(dim_red_input)*(1/3))+1)
dim_reds[["probes"]][["umap"]] = umap_hyperparam_explo(t(dim_red_input), n_neighbors=n_neighbors, min_dist=c(0.01,0.1,0.5), seeds=c(42,59,254))


```

# Plots


## Samples

```{r plot_samples, echo=T, warning=F}

#--- Variables to color samples by

samp_col_vars = names(sinfo)

#exclude any that are comments
samp_col_vars = samp_col_vars %>% .[grep("comment", ., ignore.case=T, invert=T)]

#add all probes to color by
samp_col_vars = c(samp_col_vars, colnames(mfi_input))

#--- If samples are annotated with an ELISA status, make sure all samples from same individual have the same status
if("elisa.S1" %in% names(sinfo)) {
  
  #make elisa.S1 character so that it will be plotted correctly
  sinfo$elisa.S1 = sinfo$elisa.S1 %>% as.character()
  
  #give all samples from same individual same status
  sinfo = sinfo %>% mutate(elisa.S1 = replace(elisa.S1, 
                                                subj.id %in% (sinfo %>%
                                                                    filter(elisa.S1 == "0") %>%
                                                                    .[,"subj.id"] %>%
                                                                    unlist()),
                                                "0")
                      ) %>%
                      mutate(elisa.S1 = replace(elisa.S1, 
                                                subj.id %in% (sinfo %>%
                                                                    filter(elisa.S1 == "1") %>%
                                                                    .[,"subj.id"] %>%
                                                                    unlist()),
                                                "1")
                      )
                      
  
  

}


#--- Make new variable indicating positive and negative samples
sinfo <- sinfo %>%
              mutate(subj_status = subj.id) %>%
              mutate(subj_status = replace(subj_status, subj_status == "Negative (02)", "Negative")) %>%
              mutate(subj_status = replace(subj_status, subj_status == "Negative (NR)", "Negative")) %>%
              mutate(subj_status = replace(subj_status, subj_status == "Negative (SL)", "Negative")) %>%
              mutate(subj_status = replace(subj_status, subj_status == "Positive (HD)", "Positive")) %>%
              mutate(subj_status = replace(subj_status, subj_status == "Positive (03)", "Positive")) %>%
              mutate(subj_status = replace(subj_status, grep("CBC", subj_status), "Population study")) 

if("elisa.S1" %in% names(sinfo)){
  sinfo %>% 
  mutate(subj_status = replace(subj_status, elisa.S1 == "1", "Positive")) %>%
  mutate(subj_status = replace(subj_status, elisa.S1 == "0", "Negative"))
}

samp_col_vars = c(samp_col_vars, "subj_status")

#--- Plotting

#list for storing plots
plot_list = list()

#output file for plots
pdf(file=paste0(plot_prefix, "_samples.pdf"), onefile=T, height=25, width=30)

#loop through variables to plot by
for(col_var in samp_col_vars){

  pl=list()
  
  #loop through dimensionality reduction methods
  for(method in names(dim_reds[["samples"]])){

    #loop through hyperparameter combinations (only one for PCA)
    for(hyp_comb in names(dim_reds[["samples"]][[method]])){
    
      #define coordinates and add sample info
      plot_data = dim_reds[["samples"]][[method]][[hyp_comb]] %>%
                    as_tibble() %>%
                    rename(., x = 1) %>%
                    rename(., y = 2) %>%
                    mutate(id = dim_reds[["samples"]][[method]][[hyp_comb]] %>% rownames()) %>%
                    left_join(., sinfo, by="id")
    
      #add mfi levels
      plot_data = mfi_input %>%
                    as_tibble %>%
                    mutate(id = rownames(mfi_input)) %>%
                    right_join(., plot_data, by="id")
      
      #only make plots colored by col_var if the variable contains non-NA
      if(!(plot_data[,col_var] %>% is.na() %>% all())){
      
        #create plot
        p <- ggplot(plot_data, aes_string(x="x", y="y", color=col_var, group="subj.id")) +
          geom_point(alpha=0.4) +
          labs(title = paste0(toupper(method), ", ", hyp_comb)) +
          theme_light() +
          theme(legend.position = "none", 
                plot.title=element_text(size=8)) +
          xlab("UMAP1") + ylab("UMAP2")
        
        #if variable to color by is subj_status or elisa.S1, use colors defined by global options
        if(col_var == "subj_status"){
          cols = c("Negative" = global_color_options["negatives", "color"], 
                   "Positive" = global_color_options["positives", "color"], 
                   "Population study" = global_color_options["population_study", "color"])
          
          p <- p + scale_color_manual(values = cols)
          
        } else if(col_var == "elisa.S1"){
            cols = c("0" = global_color_options["negatives", "color"], 
                    "1" = global_color_options["positives", "color"])
          
            p <- p + scale_color_manual(values = cols, na.value = "grey")
        
        #if col_var is sample_type, also make plots with lines between samples from same subject
        } else if(col_var == "subtype"){
          
          p <- p + geom_path(aes(group=subj.id, color=batch))
          
        }
        
        #extract if method is PCA
        if(method == "pca"){
          p <- p + theme(legend.position="right")
          leg <- get_legend(p)
          p <- p + theme(legend.position="none") + xlab("PC1") + ylab("PC2")
        }
        
        
        p <- ggdraw() + draw_plot(p, x = 0, y = 0, width = 1, height = 1)
        
        #save plot
        plot_name = paste(method, col_var, hyp_comb)
        plot_list[["samples"]][[plot_name]] = p
        plot_list[["samples"]][[paste0("legend ", col_var, " 1")]] = leg
        pl[[plot_name]] = p
        
      }
    }
  }
  
  #only make plots colored by col_var if the variable contains non-NA
  if(!(plot_data[,col_var] %>% is.na() %>% all())){
  
    #combine plots on page
    p <- eval(parse(text = paste0(
                        "ggarrange(", 
                        paste(
                          
                          #paste0("plot_list[['samples']][['", grep(col_var, names(plot_list[["samples"]]), value=T),"']]"
                          paste0("pl[['", names(pl),"']]"
                                 ), 
                          collapse = ", "
                        ), ", leg",
                        ", nrow=5, ncol=", ceiling(length(grep(col_var, names(plot_list[["samples"]])))/5), ", widts=1, heights=1)")
    )) +  plot_annotation(title = col_var, 
                      subtitle = footnote,
                      caption = Sys.Date(),
                      theme = theme(plot.title=element_text(size=20),
            plot.subtitle=element_text(size=10, color="grey40"),
            legend.title = element_blank()))
    
    plot(p)
  }
}

dev.off()


           
```

# Extract selected plots and corresponding classifications

For UMAP, a plot representing the collection of plots is manually selected. A threshold for defining samples as sero-positive or negative is also set manually based on visual inspection of the plot.

Note:   
Be aware that it may be necessary to update this code if input data is changed.   
The following are manually selected based on visual inspection of PCA and UMAP plots:   
- Selected representative UMAP plot   
- Cutoffs for distinguishing positive and negative individuals   

```{r extract_selected_plots, eval=T}

#output file for plots
pdf(file=paste0(plot_prefix, "_samples_selected_plots.pdf"), height=5, width=12)

#--- selection of umap hyperparameter combinations

# batch 1
if(data_set == "batch1"){
  
  if(detection_ab == "IgG"){
    neighbors = 158
    min_dist = 0.01
    seed = 254
    
    positives = dim_reds[["samples"]][["umap"]][[paste0("n_neighbors: ", neighbors, ", min_dist: ", min_dist,", seed: ", seed)]] %>% 
                    .[.[,"V2"] < -10,] %>% 
                    rownames()
    
  } else if(detection_ab == "IgM"){
    neighbors = 158
    min_dist = 0.1
    seed = 254
    
    positives = dim_reds[["samples"]][["umap"]][[paste0("n_neighbors: ", neighbors, ", min_dist: ", min_dist,", seed: ", seed)]] %>% 
                    .[.[,"V2"] > 4,] %>% 
                    rownames()
    
  } else if(detection_ab == "IgM_IgG" | detection_ab == "IgG_IgM"){
    neighbors = 25
    min_dist = 0.5
    seed = 42
    
    positives = dim_reds[["samples"]][["umap"]][[paste0("n_neighbors: ", neighbors, ", min_dist: ", min_dist,", seed: ", seed)]] %>% 
                    .[.[,"V2"] < -5,] %>% 
                    rownames()
  }
  
#batch 2
} else if(data_set == "batch2"){
    
    if(detection_ab == "IgG"){
      neighbors = 27
      min_dist = 0.5
      seed = 254
      
      positives = dim_reds[["samples"]][["umap"]][[paste0("n_neighbors: ", neighbors, ", min_dist: ", min_dist,", seed: ", seed)]] %>% 
                      .[.[,"V2"] < -5,] %>% 
                      rownames()
      
    } else if (detection_ab == "IgM"){
      neighbors = 27
      min_dist = 0.5
      seed = 59
      
      positives = dim_reds[["samples"]][["umap"]][[paste0("n_neighbors: ", neighbors, ", min_dist: ", min_dist,", seed: ", seed)]] %>% 
                      .[.[,"V1"] > 3,] %>% 
                      rownames()
      
    } else if(detection_ab == "IgM_IgG" | detection_ab == "IgG_IgM"){
      neighbors = 27
      min_dist = 0.5
      seed = 254
      
      positives = dim_reds[["samples"]][["umap"]][[paste0("n_neighbors: ", neighbors, ", min_dist: ", min_dist,", seed: ", seed)]] %>% 
                      .[.[,"V2"] < -5,] %>% 
                      rownames()
    }

#cbc03
} else if(data_set == "cbc03"){
   
  if(detection_ab == "IgG"){
      neighbors = 9
      min_dist = 0.5
      seed = 42

      positives = dim_reds[["samples"]][["umap"]][[paste0("n_neighbors: ", neighbors, ", min_dist: ", min_dist,", seed: ", seed)]] %>% 
                      .[.[,"V2"] > 0,] %>% 
                      rownames()
    }
}

#add column to sinfo with positivity status
sinfo <- sinfo %>%
          mutate(!!paste0("umap_seropositive_", detection_ab) := ifelse(.$id %in% positives, 1, 0))

# write out table with umap based sero-positive status
write.table(sinfo %>% .[,c("id","subj.id",paste0("umap_seropositive_", detection_ab))], file = paste0(plot_prefix, "_seropositivity.txt"))

#--- plotting

#loop through variables to colur by
for(col_var in samp_col_vars){
  
  #pca
  a <- plot_list[["samples"]][[paste0("pca ", col_var, " 1")]]

  #umap   
  b <- plot_list[["samples"]][[paste0("umap ", col_var, " n_neighbors: ",neighbors, ", min_dist: ", min_dist, ", seed: ", seed)]]
  
  #legend
  leg <- plot_list[["samples"]][[paste0("legend ", col_var, " 1")]]
  
  #combine plots, dont' include legend for samp
  if(col_var %in% c("id","subj.id")) {
    p <- ggarrange(a, b, nrow=1, ncol=3)  
  } else{
    p <- ggarrange(a, b, leg, nrow=1, ncol=3)  
  }
  
  #add annotation
  p <- p + plot_annotation(title = col_var, 
                      subtitle = footnote,
                      caption = Sys.Date(),
                      theme = theme(plot.title=element_text(size=20),
            plot.subtitle=element_text(size=10, color="grey40"),
            legend.title = element_blank()))
  
  plot(p)
  
  #save plot to plot in report
  if(col_var == "subj_status"){
    plot_to_show <- p
  }
  
}
  
dev.off()


```


```{r show_sample_example }
  plot(plot_to_show)
  #plot(plot_list[["samples"]][["pca sample_type 1"]])
```

## Binders

OBS: This code is not run.
No results from this code has been included in manuscript (as of 2020/07/14) and the code has not been updated recently and might therefore not work properly.

```{r plot_probes, echo=F, eval=F}

#variables to color samples by
probe_col_vars = c("Coupling_chemistry","detect_Ab","Binder")

#--- Plotting

#list for storing plots
plot_list = list()

#loop through dimensionality reduction methods
for(method in names(dim_reds[["probes"]])){
  
  #output file for plots
  pdf(file=paste0(plot_prefix, "_", method, "_probes.pdf"), onefile=T)
  
  #define coordinates
  plot_data = dim_reds[["probes"]][[method]] %>%
                as_tibble() %>%
                mutate(id = dim_reds[["probes"]][[method]] %>% rownames()) %>%
                left_join(., binder, by="id")
  
  #loop through variables to plot by
  for(col_var in probe_col_vars){
    
    p <- ggplot(plot_data, aes_string(x="PC1", y="PC2", color=col_var)) +
      geom_point(alpha=0.5) +
      labs(title = paste0(toupper(method), ", ", col_var), 
           subtitle = footnote,
           caption = Sys.Date()) +
      theme(plot.subtitle=element_text(size=7, hjust=0.5, color="grey40"),
            legend.title = element_blank())
    
    #if there is a large number of categories, put the legend under the plot and make it smaller
    if(plot_data[,col_var] %>% unique() %>% nrow() > 10){
      p <- p +
            theme(legend.position="bottom", legend.title = element_text(size=5), legend.text = element_text(size=5))

    }

    plot(p)
    
    plot_list[[paste(method, col_var, sep="_")]] = p

  }

  dev.off()   
  
         
}          


```

```{r show_binder_example, eval = F}
plot(plot_list[["pca_Binder"]])
```



# Session information

\footnotesize
```{r session_info, eval=T, echo=F}
sessionInfo()

```
\normalsize